SODA POP : Robust online Partial Ordering Planning for real life applications

### 2015-11-22 article/article.tex
Redacted Experiments + Proofs + Conclusions + sugested corrections + proof reading

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={SODA POP : Robust online Partial Ordering Planning for real life applications},
            pdfauthor={Antoine GRÉA Samir AKNINE Laetitia MATIGNON},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{listings}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

\title{SODA POP : Robust online Partial Ordering Planning for real life
applications}
\author{Antoine GRÉA Samir AKNINE Laetitia MATIGNON}
\date{}
\usepackage{xcolor, listings}
\usepackage{latex/aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{algorithm, algpseudocode}
\usepackage{amsthm}

@@

\lstset{
    basicstyle={\footnotesize\ttfamily},
    numbers=left,
    keywordstyle=\color{solarized-orange}\bfseries,
%    identifierstyle=\color{solarized-base02},
    stringstyle=\color{solarized-cyan},
    commentstyle=\color{solarized-base1}\itshape,
    numberstyle=\tiny\color{solarized-base0},
    columns=flexible,
    stepnumber=1,
    numbersep=5pt,
    backgroundcolor=\color{solarized-base3},
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2,
    captionpos=b,
    breaklines=true,
    breakatwhitespace=true,
    breakautoindent=true,
    escapeinside={\%*}{*)},
    linewidth=\textwidth,
    basewidth=0.5em,
}

@@

\begin{document}
\maketitle
\begin{abstract}
In cognitive applications the needs of ordering action to reach a goal
created the dommain of automated planning. This domain is actually
domined by researches arround performance and the advances in state
based planning has eclipsed other less efficient ways of doing. Partial
Ordering Planning (POP) is one of the most flexible way to order actions
and create plan that can be refined easily. This can be used to achieve
greater resilience, repairing capabilities and soft resolution that
applications like online plan recognision or decision making can benefit
from. This calls for new solutions that our set of algorithm named Soft
Ordering and Defect Aware Partial Ordering Planning (SODA POP) targets
with plan quality in mind. The algorithm seeks a new set of defects in
broken input and actions in order to fix them and even to derive the
problem in order to always output a valid plan in case the problem is
imposible to solve.
\end{abstract}

@@

For some time Partial Order Planning (POP) has been the most popular
approach to planning resolution. This kind of algorithms are based on
\emph{least commitment strategy} on plan step ordering that can allow
actions to be flexibly interleaved during execution {[}1{]}. Thus the
way the search is made using flexible partial plan as a search space
allowed for more versatility for a wide variety of uses. As of more
recent years, new state space search models and heuristics {[}2{]} have
been demonstrated to be more efficient than POP planners due to the
simplicity and relevance of states representation opposed to partial
plans {[}3{]}. This have made the search of performance the main axis of
research for planning in the community.

While this goal is justified, it shadows other problems that some
practical applications cause. For example, the flexibility of Plan Space
Planning (PSP) is an obvious advantage for applications needing online
planning: plans can be repaired on the fly as new informations and
objectives enter the system. The idea of using POP for online planning
and repairing plans instead of replanning everything is not new {[}4{]},
but has never been paired with the resilience that some other cognitive
applications may need, especially when dealing with sensors data and
noise.

This resilience makes fixing errors easier than with an external
algorithm as the plan logic allows for context driven decision on the
way to repair the issues. For example if an action becomes unrelevant or
incoherent the flaw in the partial plan makes the problem to fix
explicit and therefore easier to fix. The client softwares might
sometimes provide plans that can contain errors and imperfections that
will decrease significantly the efficiency of the computation and the
quality of the output plan.

Adding to that, these plans may become totally unsolvable. This problem
is to our knowledge not treated in planning of all forms (state
planning, PSP, and even constraint satisfaction planning) as usually the
aim is to find a solution relative to the original plan (which makes
sense). But as we proceed a mechanism of \emph{problem} fixing may be
required. This will allow soft solving of any problem regardless of its
solvability.

One of the application that needs these improvements is plan recognition
with the particular use of off-the-shelf planners to infer the pursued
goal of an agent where online planning and resilience is particularly
important.

These problems call for new ways to improve the answer and behavior of a
planner. These improvements must provide relevant plan information
pointing out exactly what needs to be done in order to make a planning
problem solvable, even in the case of obviously broken input data. We
aim to solve this in this paper while preserving the advantages of PSP
algorithms (flexibly, easy fixing of plans, soudness and completeness).
Our Soft Ordering and Defect Aware Partial Ordering Planning (SODA POP)
algorithm will target those issues.

Our new set of auxiliary algorithms allows to make POP algorithm more
resilient, efficient and relevant. This is done by pre-emptively
computing proper plans for goals, by solving new kinds of defects that
input plans may provide, and by healing compromised plan by extending
the initial problem to allow resolution.

To explain this work we first describe a few notions, notations and
specificities of existing POP algorithms. Then we present and illustrate
their limitations, categorising the different defects arising from the
resilience problem and explaining our complementary algorithms, their
uses and properties. To finish we compare the performance, resilence and
quality of POP and our solution.

@@

In order to present our work and explain examples we introduce a way of
representation for schema in figure \ref{fig:legend} and notations for
mathematical representation.

\begin{figure}[htbp]
\centering
\includegraphics{graphics/legend.pdf}
\caption{Global legend for how partial plans are represented in the
current paper\label{fig:legend}}
\end{figure}

\subsection{Classic planning}\label{classic-planning}

@@

A fluent is a property of the world. It is often represented by first
order logical propositions but in our case we choose to focus on the
algorithm and to represent fluents as simple literals (fully
instantiated). We note \(\lnot f\) the complementary fluent of \(f\)
meaning that \(f\) is true when \(\lnot f\) is false and vice-versa.

In order to make our example simpler we use \(\mathbb{Q}^*\), the set of
relative integers without \(0\), as the fluent domain. We use negative
integers to represent opposite fluents.

@@

@@

We define a state as a set of fluents. States can be additively
combined. We note
\(s_1 + s_2 = \left( s_1 \cup s_2 \right) - \left\{ f \middle| f \in s_1 \land \lnot f \in s_2 \right\}\)
such operation. It is the union of the fluents with an erasure of the
complementary ones.

@@

An action \(a\) can be functionally applied to a state \(s\) following :
\[a:= \substack{ \left\{ s \models a \middle| s \in S\right\} \to S\\
    a(s) \mapsto s + eff(a)}\] with \(S\) being the set of all states.
There are some special names for actions. An action with no
preconditions is synonymous to a state and one with empty effect is
called a goal.

@@

@@

A \emph{partial plan} is a tuple \(p = \langle A_p, L\rangle\) where
\(A_p\) is a set of steps (actions) and \(L\) is a set of causal links
of the form \(a_i \xrightarrow{f} a_j\), such that
\(\{ a_i, a_j \} \subset A_p \land f \in eff(a_i) \cap pre(a_j)\) or
said otherwise that \(f\) is provided by \(a_i\) to \(a_j\) via this
causal link. We include the ordering constraints of PSP in the causal
links. An ordering constraint is noted \(a_i \to a_j\) and means that
the plan consider \(a_i\) as a step that is prior to \(a_j\) without
specific cause (usually because of threat resolution).

@@

@@

When refining a partial plan, we need to fix flaws. Those could be
present or be created by the refining process. Flaws can either be
unsatisfied subgoals or threats to causal links.

@@

A resolver for a subgoal is an action \(a_r \in A\) that has \(s\) as an
effect \(s \in eff(a_r)\). It is inserted along with a causal link noted
\(a_r \xrightarrow{s} a_s\).

@@

A step \(a_t\) is said to threaten a causal link
\(a_i \xrightarrow{t} a_j\) if and only if
\[\neg t \in eff(a_t) \land a_i \succ a_t \succ a_j \models L\] Said
otherwise, the action has a possible complementary effect that can be
inserted between two actions needing this fluent while being consistant
with the ordering constraint in \(L\).

The usual resolvers are either \(a_t \to a_i\) or \(a_j \to a_t\) which
are called respectively promotion and demotion links. Another resolver
is called a white knight that is an action \(a_k\) that reestablish
\(t\) after \(a_t\).

@@

@@

Partial Order Planning (POP) is a popular implementation of the general
PSP algorithm. It is proven to be sound and complete {[}5{]}. The
completeness of the algorithm guarantees that if the problem has a
solution it will be found by the algorithm. The soundness assures that
any answer from the algorithm is valid. POP refines a partial plan by
trying to fix its flaws.

@@

\begin{algorithm}\caption{Classical Partial Order Planning algorithm}\label{pop}\begin{algorithmic}

@@

This standard way of doing have seen multiple improvements over
expressiveness like with UCPOP {[}6{]}, hierarchical task network to add
more user control over sub-plans {[}7{]}, cognition with defeasible
reasoning {[}8{]}, or speed with multiple ways to implement the popular
fast forward method from state planning {[}9{]}. However, all these
variants do not treat the problem of online planning, resilience and
soft solving.

Some other closer works like {[}4{]} treats the problem of online
planning by removing big chuncks of the partial plan by identifying
incorect trees in the plan. This causes an heavy replanning of the
problem even if only one action needed removal. This is a big problem
when trying to adapt a plan with minimal changes due to replanning.

@@

Before continuing, we present a simple example of classical POP
execution with the problem represented in figure \ref{fig:problem}. We
did not represent empty preconditions or effects to improve readability.
Here we have an initial state
\(I = \langle \emptyset , \{ 1, 2 \} \rangle\) and a goal
\(G = \langle \{ 3, 4, -5, 6 \}, \emptyset \rangle\) encoded as dummy
steps. We also introduce actions that are not steps yet but that are
provided by \(A\). The actions \(a\), \(b\) and \(c\) are normal actions
that are useful to achieve the goal. The action \(t\) is meant to be
threatening to the plan's integrity and will generate threats. We
introduce \(u\) as a useless action, \(v\) as a toxic action, \(w\) as a
dead-end action and \(x\) as a contradictory action.

@@

This example has been crafted to illustrate problems with standard POP
implementations. We give a possible resulting plan of standard POP in
figure \ref{fig:pop}. We can see some issues as for how the plan has
been built. The action \(v\) is being used even if it is useless since
\(b\) already provided fluent \(4\). We can also notice that despite
being contradictory the action \(x\) raised no alarm. As for ordering
constraints we can clearly see that the link \(a \to t\) is redundant
with the path \(a \xrightarrow{5} c \to t\) that already puts that
constraint by transitivity. Also some problems arise during execution
with the selection of \(w\) that causes a dead-end.

@@

In order to improve POP algorithms' resilience, online performance and
plan quality, we propose a set of auxiliary algorithms that provides POP
with a clean and efficiently populated initial plan.

@@

\begin{algorithm}\caption{Proper plan generation algorithm for a given goal $g$}\label{properplan}\begin{algorithmic}

@@

On our example it illustrate really well the need for resilience. The
algorithm simply recursively chose the actions and end up with the
partial plan represented in figure \ref{fig:proper}. The partial plan
doesn't have an initial state (because of its offline nature). This plan
also shows several cycles and obvious problems. Moreover the plan has
all the steps of the correct final plan and therefore remove the subgoal
search in POP if it uses this plan has input (after it has been cleaned
obviously).

@@

@@

\begin{algorithm}\caption{Defect resolution algorithm}\label{defectresolution}\begin{algorithmic}

\Function{clean}{Problem $P$} \State \Call{illegal}{$P$}
\State \Call{interfering}{$P$} \EndFunction

\Function{illegal}{Problem $P$}
\State \Call{assert}{$pre(P.I) = \emptyset$}
\State \Call{assert}{$eff(P.G) = \emptyset$}
\State \(P.p.A_p \gets P.p.A_p \cup \{P.I, P.G\}\)
\State \Call{breakCycles}{$P$} \State Actions
\(actions \gets P.A \cup P.p.A_p\) \ForAll{Action $a \in actions$}
\State \Call{inconsistent}{$a$, $P$} \State \Call{toxic}{$a$, $P$}
\EndFor
    \State \Call{liarLinks}{$P$} \EndFunction

\Function{interfering}{Problem $P$} \State \Call{repeating}{$P$}
\State \Call{useless}{$P$} \EndFunction

@@

When the POP algorithm is used to refine a given plan (that was not
generated with POP or that was altered), a set of new defects can be
present in it interfering in the resolution and sometimes making it
impossible to solve. We emphasize that these defects are not regular POP
flaws but new problems that classical POP can't solve. The aim of this
auxiliary algorithm is to clean the plans from such defects in order to
improve computational time, resilience and plan quality . It should be
noted that in some cases cleaning plans will increase the number of
flaws in the plan but will always improve the overall quality of it.

@@

\begin{definition}[Illegal defects]

These defects are usually hypothesized out by regular models. They are
illegal use of partial plan representation and shouldn't happen under
regular circumstances. They may appear if the input is generated by an
approximate cognitive model that doesn't ensure consistency of the
output or by unchecked corrupted data.

@@

A plan cannot contain cycles as it makes it impossible to complete.
Cycles are usually detected as they are inserted in a plan but poor
input can potentially contains them and breaks the POP algorithm as it
cannot undo cycles.

We use a popular and simple Depth First Search (DFS) algorithm to detect
cycles. Upon cycle detection the algorithm can remove arbitrarily a link
in the cycle to break it. The most effective solution is to remove the
link that is the farthest from the goal travelling backward as it would
be that link that would have been last added in the regular POP
algorithm.

>>>>

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{unicode=true,
            pdftitle={SODA POP : Robust online Partial Ordering Planning for real life applications},
            pdfauthor={Paper ID\#232},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{listings}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}

\title{SODA POP : Robust online Partial Ordering Planning for real life
applications}
\author{Paper ID\#232}
\date{}
\usepackage{xcolor, listings}
\usepackage{latex/aaai}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{amsthm}

@@

%%%%%% for algorithms
\errorcontextlines\maxdimen
\algrenewcommand\alglinenumber[1]{\tiny\color{solarized-base0} #1}

% begin vertical rule patch for algorithmicx (http://tex.stackexchange.com/questions/144840/vertical-loop-block-lines-in-algorithmicx-with-noend-option)
\makeatletter
% start with some helper code
% This is the vertical rule that is inserted
    \newcommand*{\algrule}[1][\algorithmicindent]{\makebox[#1][l]{\color{solarized-base0}\hspace*{.5em}\thealgruleextra\vrule height \thealgruleheight depth \thealgruledepth}}%
% its height and depth need to be adjustable
\newcommand*{\thealgruleextra}{}
\newcommand*{\thealgruleheight}{.75\baselineskip}
\newcommand*{\thealgruledepth}{.25\baselineskip}

\newcount\ALG@printindent@tempcnta
\def\ALG@printindent{%
    \ifnum \theALG@nested>0% is there anything to print
        \ifx\ALG@text\ALG@x@notext% is this an end group without any text?
            % do nothing
        \else
            \unskip
            \addvspace{-1pt}% FUDGE to make the rules line up
            % draw a rule for each indent level
            \ALG@printindent@tempcnta=1
            \loop
                \algrule[\csname ALG@ind@\the\ALG@printindent@tempcnta\endcsname]%
                \advance \ALG@printindent@tempcnta 1
            \ifnum \ALG@printindent@tempcnta<\numexpr\theALG@nested+1\relax% can't do <=, so add one to RHS and use < instead
            \repeat
        \fi
    \fi
    }%
\usepackage{etoolbox}
% the following line injects our new indent handling code in place of the default spacing
\patchcmd{\ALG@doentity}{\noindent\hskip\ALG@tlm}{\ALG@printindent}{}{\errmessage{failed to patch}}
\makeatother

% the required height and depth are set by measuring the content to be shown
% this means that the content is processed twice
\newbox\statebox
\newcommand{\myState}[1]{%
    \setbox\statebox=\vbox{#1}%
    \edef\thealgruleheight{\dimexpr \the\ht\statebox+1pt\relax}%
    \edef\thealgruledepth{\dimexpr \the\dp\statebox+1pt\relax}%
    \ifdim\thealgruleheight<.75\baselineskip
        \def\thealgruleheight{\dimexpr .75\baselineskip+1pt\relax}%
    \fi
    \ifdim\thealgruledepth<.25\baselineskip
        \def\thealgruledepth{\dimexpr .25\baselineskip+1pt\relax}%
    \fi
    %\showboxdepth=100
    %\showboxbreadth=100
    %\showbox\statebox
    \State #1%
    %\State \usebox\statebox
    %\State \unvbox\statebox
    %reset in case the next command is not wrapped in \myState
    \def\thealgruleheight{\dimexpr .75\baselineskip+1pt\relax}%
    \def\thealgruledepth{\dimexpr .25\baselineskip+1pt\relax}%
}
% end vertical rule patch for algorithmicx

%%%% For listings (code blocks)
\lstset{
    basicstyle={\footnotesize\ttfamily},
    numbers=left,
    keywordstyle=\color{solarized-orange}\bfseries,
%    identifierstyle=\color{solarized-base02},
    stringstyle=\color{solarized-cyan},
    commentstyle=\color{solarized-base1}\itshape,
    numberstyle=\tiny\color{solarized-base0},
    columns=flexible,
    stepnumber=1,
    numbersep=5pt,
    backgroundcolor=\color{solarized-base3},
    showspaces=false,
    showstringspaces=false,
    showtabs=false,
    tabsize=2,
    captionpos=b,
    breaklines=true,
    breakatwhitespace=true,
    breakautoindent=true,
    escapeinside={\%*}{*)},
    linewidth=\textwidth,
    basewidth=0.5em,
}

%%%% For links URL
\hypersetup{colorlinks=true, linkcolor=solarized-blue, citecolor=solarized-base1, filecolor=solarized-magenta, urlcolor=solarized-cyan}

@@

\begin{document}
\maketitle
\begin{abstract}
As of recent years, automated planning domain has mainly focused on
performances and advances in state space planning to improve
scalability. That orientation shadows other less efficient ways of doing
like Partial Ordering Planning (POP) that has the advantage to be much
more flexible. This approach \emph{generates} plans that can be refined
easily and can be used to achieve greater resilience, repairing
capabilities and soft resolution that applications \emph{based on}
online plan recognition or decision-making can benefit from. \emph{This
paper presents} a set of algorithms, named Soft Ordering and Defect
Aware Partial Ordering Planning (SODA POP), that aims at targeting these
objectives and maintain a high plan quality. Our algorithms create
proper offline plans for goals and use an effective defect fixing
\emph{algorithm} to repair input plans even if these plans are
corrupted. SODA POP can also use healer actions and links in order to
always return a valid plan even in cases where the problem is impossible
to solve using problem derivation. Some relevant properties of these
algorithms are analyzed in this article and experimental results show
interesting performances for online planning and repairing.
\end{abstract}

@@

For some time Partial Order Planning (POP) has been the most popular
approach to planning resolution. This kind of algorithms are based on
\emph{least commitment strategy} on plan step ordering that can allow
actions to be flexibly interleaved during execution {[}1{]}. Thus, the
way the search is made using flexible partial plan as a search space
allowed for more versatility for a wide variety of uses. As of more
recent years, new state space search models and heuristics {[}2{]} have
been demonstrated to be more efficient than POP planners due to the
simplicity and relevance of states representation opposed to partial
plans {[}3{]}. This have made the search of performance the main axis of
research for planning in the community .

While this goal is justified, it shadows other problems that some
practical applications cause. For example, the flexibility of Plan Space
Planning (PSP) is an obvious advantage for applications needing online
planning: plans can be repaired on the fly as new informations and
objectives enter the system. The idea of using POP for online planning
and repairing plans instead of replanning everything is not new {[}4{]},
but has never been paired with the resilience that some other cognitive
applications may need, especially when dealing with sensors noise in
input data.

This resilience makes fixing errors easier than with an external
algorithm as the plan logic allows for context driven decision on the
way to repair the issues. For example if an action becomes unrelevant or
incoherent, the flaw in the partial plan makes the issue explicit and
therefore easier to fix. Softwares might sometimes provide plans that
can contain errors and imperfections that will decrease significantly
the efficiency of the computation and the quality of the output plan.

Adding to that, these plans may become totally unsolvable. This problem
is to our knowledge not treated in planning of all forms (state
planning, PSP, and even constraint satisfaction planning) as usually the
aim is to find a solution relative to the original plan (which makes
sense). But as we proceed a mechanism of \emph{problem derivation} may
be required. This will allow soft solving of any problem regardless of
its resolvability.

One of the applications that needs these improvements is plan
recognition with the particular use of off-the-shelf planners to infer
the pursued goal of an agent where online planning and resilience is
particularly important {[}5{]}. This method adds dummy actions that need
to be satisfied by modifying the goal to ensure that the observed plan
is selected by the planner. A system that is able to repair plan in real
time will ease such an application. Another application is
decision-making in dynamical environments. Indeed, having a plan that
details all steps and explicits the ones that are not possible, can help
decide of the line of action to take.

These problems call for new ways to improve the answer of a planner.
These improvements must provide relevant plan information pointing out
exactly what needs to be done in order to make a planning problem
solvable, even in the case of obviously inconsistent input data. This
paper aims to solve this while preserving the advantages of PSP
algorithms (flexibility, easy fixing of plans, soundness and
completeness). Our Soft Ordering and Defect Aware Partial Ordering
Planning (SODA POP) algorithm will target those issues.

Our new set of auxiliary algorithms allows making POP algorithm more
resilient, efficient and relevant. This is done by pre-emptively
computing proper plans for goals, by solving new kinds of defects that
input plans may exhibit, and by healing compromised plan by deriving the
initial problem with forged actions to allow resolution.

To explain this work we first introduce a few notions, notations and
specificities of existing POP algorithms. Then we present and illustrate
their limitations, categorizing the different defects arising from the
resilience problem and explaining our auxiliary algorithms, their uses
and properties. To finish we compare the performance, resilience and
quality of POP and our solution.

@@

First we introduce some definitions and notations for mathematical
representation to present our model.

\subsection{Classical planning}\label{classical-planning}

@@

A fluent is a property of the world. We note \(\lnot f\) the
complementary fluent of \(f\) meaning that \(f\) is true when
\(\lnot f\) is false and vice-versa.

@@

It is often represented by first order logical propositions but in this
paper we choose to focus on the algorithm and to represent fluents as
simple literals (fully instantiated) using \(\mathbb{Q}^*\), the set of
relative integers without \(0\), as the fluent domain. We use negative
integers to represent opposite fluents.

@@

A state is defined as a set of fluents. States can be additively
combined. We note
\(s_1 + s_2 = \left( s_1 \cup s_2 \right) - \left\{ f \middle| f \in s_1 \land \lnot f \in s_2 \right\}\)
such operation. It is the union of the fluents with an erasure of the
complementary ones.

@@

An action \(a\) can be functionally applied to a state \(s\) as follows
: \[a:= \substack{ \left\{ s \models a \middle| s \in S\right\} \to S\\
    a(s) \mapsto s + eff(a)}\] with \(S\) being the set of all states.
This means that an action adds all its effect to the state of the world
on application (by removing complementary fluents if needed)

@@

We distinguish between two specific kinds of actions : actions with no
preconditions are synonymous to a state and those with empty effect are
called a goal.

@@

A \emph{partial plan} is a tuple \(p = \langle A_p, L\rangle\) where
\(A_p\) is a set of steps (actions) and \(L\) is a set of causal links
of the form \(a_i \xrightarrow{f} a_j\), such that
\(\{ a_i, a_j \} \subset A_p \land f \in eff(a_i) \cap pre(a_j)\) or
said otherwise, this causal link means that the fluent \(f\) is provided
by an effect of \(a_i\) to a precondition of \(a_j\). We include the
ordering constraints of PSP in the causal links. An ordering constraint
is noted \(a_i \to a_j\) and means that the plan means \(a_i\) as a step
that is prior to \(a_j\) without specific cause (usually because of
threat resolution). We note the order relation over \(A\) for an action
\(a_i\) that is prior to the action \(a_j\) following all the causal
links and ordering constraints \(a_i \succ a_j\) .

@@

The figure \ref{fig:legend} details how the elements of partial plans
are represented in the rest of this paper.

\begin{figure}[htbp]
\centering
\includegraphics{graphics/legend.pdf}
\caption{Global legend for how partial plans are represented in this
paper\label{fig:legend}}
\end{figure}

In this figure the fluent \(5\) is a precondition of the action \(a\)
and the fluent \(6\) an effect. The causal link is providing the fluent
\(7\) (only as a representation example or it would be a liar link). The
dotted line is an ordering constraint that would mean that \(a\) should
be placed before \(a\) (for representation still or it would be a
cycle). \(I\) and \(G\) are respectively the initial and goal step.

@@

When refining a partial plan, we need to fix flaws. Those could be
present in the input or created by the refining process. Flaws can
either be unsatisfied subgoals or threats to causal links.

@@

@@

A step \(a_t\) is said to threaten a causal link
\(a_i \xrightarrow{t} a_j\) if and only if
\[\neg t \in eff(a_t) \land a_i \succ a_t \succ a_j \models L\]

@@

In other words, the action has a possible complementary effect that can
be inserted between two actions needing this fluent while being
consistent with the ordering constraint in \(L\).

\subsubsection{Resolvers}\label{resolvers}

Resolvers are a set of actions and causal links
\(r = \langle A_r , L_r \rangle\) that fixes flaws. A resolver for a
subgoal is an action \(a_r \in A\) that has \(s\) as an effect
\(s \in eff(a_r)\) inserted along with a causal link noted
\(a_r \xrightarrow{s} a_s\). The usual resolvers for a threat are either
\(a_t \to a_i\) or \(a_j \to a_t\) which are called respectively
promotion and demotion links. Another resolver is called a white knight
that is an action \(a_k\) that reestablishes \(t\) after \(a_t\) along
with an ordering constraint \(a_t \to a_k\) and a causal link
\(a_k \xrightarrow{t} a_j\).

@@

Partial Order Planning (POP) is a popular implementation of the general
PSP algorithm. It is proven to be sound and complete {[}6{]}. The
completeness of the algorithm guarantees that if the problem has a
solution it will be found by the algorithm. The soundness assures that
any answer from the algorithm is valid. POP refines a partial plan by
trying to fix its flaws.

@@

\begin{algorithm}\caption{Classical Partial Order Planning algorithm}\label{pop}\begin{algorithmic}[1]

@@

This standard way of doing have seen multiple improvements over
expressiveness like with UCPOP {[}7{]}, hierarchical task network to add
more user control over sub-plans {[}8{]}, cognition with defeasible
reasoning {[}9{]}, or speed with multiple ways to implement the popular
fast forward method from state planning {[}10{]}. However, all these
variants do not treat the problem of online planning, resilience and
soft solving.

Some other closer works like {[}4{]} treats the problem of online
planning by removing big chunks of the partial plan by identifying
incorrect trees in the plan. This along with heuristic choice of
unrefinemment strategies. This causes a heavy replanning of the problem
even if only one action needed removal. This also lead to an exponential
number of plan to consider. This is a big problem when trying to adapt a
plan with minimal changes due to replanning.

@@

Before continuing, we present a simple example of classical POP
execution with the problem represented in figure \ref{fig:problem}. We
did not represent empty preconditions or effects to improve readability.
Here we have an initial state
\(I = \langle \emptyset , \{ 1, 2 \} \rangle\) and a goal
\(G = \langle \{ 3, 4, -5, 6 \}, \emptyset \rangle\) encoded as dummy
steps. We also introduce actions that are not steps yet but that are
provided by \(A\). The actions \(a\), \(b\) and \(c\) are normal actions
that are useful to achieve the goal. The action \(t\) is meant to be
threatening to the plan's integrity and will generate threats. We
introduce \(u\) and \(v\) as a toxic action, \(w\) as a dead-end action
and \(x\) as a contradictory action. These notions will be defined
\protect\hyperlink{defects}{later}.

@@

This example has been crafted to illustrate problems with standard POP
implementations. We give a possible resulting plan of standard POP in
figure \ref{fig:pop}. We can see some issues as for how the plan has
been built. The action \(v\) is being used even if it is useless since
\(b\) already provided fluent \(4\). We can also notice that despite
being contradictory the action \(x\) raised no alarm. As for ordering
constraints, we can clearly see that the link \(a \to t\) is redundant
with the path \(a \xrightarrow{5} c \to t\) that already puts that
constraint by transitivity. Also, some problems arise during execution
with the selection of \(w\) that causes a dead-end.

@@

In order to improve POP algorithms' resilience, online performance and
plan quality, we propose a set of auxiliary algorithms that provides POP
with a clean and efficiently populated initial plan. The complete
algorithm will be presented in \protect\hyperlink{soda}{the next
section} as a combination of all auxiliary algorithms and regular POP.

@@

\begin{algorithm}\caption{Proper plan generation algorithm for a given goal $g$}\label{properplan}\begin{algorithmic}[1]

@@

@@

On our example it illustrates really well the need for resilience. The
algorithm simply recursively chose the actions and end up with the
partial plan represented in figure \ref{fig:proper}. The partial plan
doesn't have an initial state (because of its offline nature). This plan
also shows several cycles and obvious problems. Moreover, the plan has
all the steps of the correct final plan and therefore remove the subgoal
search in POP if it uses this plan has input (after it has been cleaned
obviously).

@@

\begin{algorithm}\caption{Defect resolution algorithm}\label{defectresolution}\begin{algorithmic}[1]

\Function{clean}{Problem $P$}
\State \Call{assert}{$pre(P.I) = \emptyset$}
\State \Call{assert}{$eff(P.G) = \emptyset$}
\State \(P.p.A_p \gets P.p.A_p \cup \{P.I, P.G\}\)
\State \(defects \gets\) \Call{findDefects}{$P$}
\State \Call{fix}{$defects$, $P$} \EndFunction

\Function{fix}{Defects $defects$, Problem $P$}
\ForAll{Defect $d \in defects$} \State \(defect.\)\Call{fix}{}
\Comment{Fixing is proper to the defect} \EndFor
\EndFunction

\Function{findDefects}{Problem $P$} \State \Return \Call{illegal}{$P$}
\(+\) \Call{interfering}{$P$} \EndFunction

@@

When the POP algorithm is used to refine a given plan (that was not
generated with POP or that was altered), a set of new defects can be
present in it interfering in the resolution and sometimes making it
impossible to solve. We emphasize that these defects are not regular POP
flaws but new problems that classical POP can't solve. The aim of this
auxiliary algorithm is to clean the plans from such defects in order to
improve computational time, resilience and plan quality. It should be
noted that in some cases cleaning plans will increase the number of
flaws in the plan but will always improve the overall quality of it.

\begin{definition}[Plan Quality]

Plan quality is an indicator that is measured by the number of defects
in a partial plan. There are two specific indicators for a plan quality
: the action quality and the link quality. We define the link and action
quality as respectively the number of link or action related defects
divided by the number of links or actions in the resulting plan.
\[quality_{action}(p) =  1 - \left( \#ActionDefects(p) \over \#p.A_p \right)\]

From there it is obvious that plan quality will improve over POP since
the present algorithm is guaranteed to remove all defects in a plan.

@@

\subsubsection{Illegal defects}\label{illegal-defects}

These defects are usually hypothesized out by regular models. They are
illegal use of partial plan representation and shouldn't happen under
regular circumstances. They may appear if the input is generated by an
approximate cognitive model that doesn't ensure consistency of the
output or by unchecked corrupted data. These defects will most of the
time simply break regular POP algorithms (from existing cycles) or at
least make the performances decrease significantly (from toxic actions).

A plan cannot contain cycles as it makes it impossible to complete.
Cycles are usually detected as they are inserted in a plan but poor
input can potentially contain them and breaks the POP algorithm as it
cannot undo cycles.

@@

We use strongly connected component detection algorithm to detect
cycles. Upon cycle detection the algorithm can remove arbitrarily a link
in the cycle to break it. The most effective solution is to remove the
link that is the farthest from the goal traveling backward as it would
be that link that would have been last added in the regular POP
algorithm. But finding this is extremely time consuming and most of the
time the cycles are of one or two actions so we choose to keep that
algorithm simpler and faster.
